Standardization of ceramic shape: A case study from the Iron Age pottery from northeastern Taiwan

### 2020-08-22 analysis/paper/paper.Rmd
edit discussion section

The shape variation reported here is subtle and invites consideration of another possible scenario, namely that the absence of major changes in vessel shape at Kiwulan may have been an act to show ethnic identity when experiencing foreign influences [cf. @Torrence2000]. Ceramic morphology could be a signal for communication between potters, such that repeating the same shape or slight modifications may have occurred in a non-deterministic way [@Kubler1962]. As the only type of locally made pottery throughout 600 years, the homogeneous and even more standardized shape in the Chinese period might convey some meaningful information about community identity. We recognize the decline in population may lead to a standardized shape, however, the evidence of ceramic spatial distribution shows a dispersed or random pattern across the whole area throughout three phases. This indicates manufacturing activities may be limited to the household scale and reflect a common, shared practice in the community. The standardization in vessel shape over time draws our attention to the endurance of traditional pottery production practices amid intrusions from Europeans and Chinese. In a culture contact situation, we speculate that social identity might have be expressed through material practices as a means of expressing cultural homogeneity and distinction from other groups [@Voss2005].

It is also important to recognize that social identity might be more complicated in a colonial context, and maybe representative of more than a colonized–colonizer or local/foreign dichotomy [@Voss2005; @Voss2008]. For example, the Shamaoshan cemetery (BC 250 to 55 AD) in Southwest China suggests that the process of the incorporation of Southwest China into the Han Empire involved a century of conflicts, resistance, and acceptance among social groups with different identities, especially in the historical context of Han immigrants [@Wu2019]. Another example demonstrates Indigenous endurance and resilience through the long-term  persistence of mound use in Mississippian culture after contact with European colonizers [@Holland]. A similar dynamic may have occurred at Kiwulan, with vessel shape indicating both acceptance of foreign influence through increased shape standardization, and resistance through the overall continuity in vessel shape. Vessel shape may be viewed as a symbolic expression of Indigenous identity and social boundaries because shape is a highly visible trait compared with other features of pottery [cf. @Roux2015]. Although there is an increase in number of imported ceramics through time at Kiwulan, production of local ceramics was continuous, and increasingly standardized. This might imply not only the utilitarian function, but an intentional and increased emphasis on the local ceramic tradition, their cultural custom, as a response to intensified foreign contact [cf. @Acabado2017]. However, additional lines of evidence are necessary from Kiwulan to confirm this speculation.

>>>>

The shape variation reported here is subtle and invites consideration of another possible scenario, namely that the absence of major changes in vessel shape at Kiwulan may have been an act to show ethnic identity when experiencing foreign influences [cf. @Torrence2000]. Ceramic morphology could be a signal for communication between potters, such that repeating the same shape or slight modifications may have occurred in a non-deterministic way [@Kubler1962]. As the only type of locally made pottery throughout 600 years, the homogeneous and even more standardized shape in the Chinese period might convey some meaningful information about community identity. We recognize the decline in population may lead to a standardized shape, however, the evidence of ceramic spatial distribution shows a dispersed or random pattern across the whole area throughout three phases. This indicates manufacturing activities may be limited to the household scale and reflect a common, shared practice in the community. The standardization in vessel shape over time draws our attention to the endurance of traditional pottery production practices amid intrusions from Europeans and Chinese. In a culture contact situation, we speculate that social identity might have be expressed through material practices as a means of expressing cultural homogeneity and distinction from other groups [@Voss2005].  
It is also important to recognize that social identity might be more complicated in a colonial context, and maybe representative of more than a colonized–colonizer or local/foreign dichotomy [@Voss2005; @Voss2008]. For example, the Shamaoshan cemetery (BC 250 to 55 AD) in Southwest China suggests that the process of the incorporation of Southwest China into the Han Empire involved a century of conflicts, resistance, and acceptance among social groups with different identities, especially in the historical context of Han immigrants [@Wu2019]. At the Oconee Valley (AD 1540 to 1670) in the Southeastern United States, Indigenous endurance and resilience is indicated by the long-term persistence of mound use lasting for 130 years after contact with European colonizers [@Holland]. A similar dynamic may have occurred at Kiwulan, with vessel shape indicating both acceptance of foreign influence through increased shape standardization, and resistance through the overall continuity in vessel shape. Vessel shape may be viewed as a symbolic expression of Indigenous identity and social boundaries because shape is a highly visible trait compared with other features of pottery [cf. @Roux2015]. Although there is an increase in number of imported ceramics through time at Kiwulan, production of local ceramics was continuous, and increasingly standardized. This might imply not only the utilitarian function, but an intentional and increased emphasis on the local ceramic tradition, their cultural custom, as a response to intensified foreign contact [cf. @Acabado2017]. However, additional lines of evidence are necessary from Kiwulan to confirm this speculation.
